Intensity Modulated Radiation Therapy using Tomotherapy

### 2014-12-09 BME 560/Paper.tex
Almost done

\documentclass[10pt,journal,compsoc]{IEEEtran} %draftclsnofoot, 
\usepackage{graphicx}
\graphicspath{ {./_figures/} }
\setlength\fboxsep{1pt}
\setlength\fboxrule{1pt}
\usepackage{multicol}
\usepackage{bigstrut}
\usepackage{tabularx}
\usepackage{tabulary}
\usepackage{booktabs}
\usepackage{amsmath}
\setlength{\parindent}{0em}
\setlength{\parskip}{1em}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{hyperref}
\hypersetup{
  colorlinks = false,
  hidelinks = true
	}
\usepackage[english]{babel}
\usepackage{blindtext}
\usepackage{times}
\usepackage{cite}

@@

  The setup has a 
  linear accelerator mounted on a rotating gantry as seen in CT 
  sacnners. The radiation is delivered to a patient in a helical way, obtained 
  by concurrent gantry rotation and couch/patient travel as shown in Fig. 
  \ref{fig1}. In general, the helical tomotherapy units have significantly 
  different radiation fields when compared to other radiotherapy systems. This 
  difference can be owed to the absence of absence of a flattening filter, a 
  thin target, an electron stopper, a beam hardener and a compact primary 
  collimator. This paper will briefly highlight the 
  radiation characteristics of the helical tomotherapy unit, especially in 
  parts where it differs from standard accelerators.

@@

      \toprule[0.3mm]

@@

      \midrule

@@

  Figure \ref{radcharb} shows the differences in the photon spectrum between 
  the 
  treatment and imaging modes of helical tomotherapy. Notice that the in the 
  imaging mode the spectrum has a lower energy when compared to treatment mode; 
  this is done so as to obatin better images by improving the contrast. A more 
  comprehensive comparision is documented in Ref. \cite{Jeraj2004}

@@

      \label{tab:compare}%
    \end{table*}%
  Tomotherapy is different from other standard IMRT techiniques such as 
  conformal radiotherapy (CRT) techniques and non-standard IMRT techniques like 
  Cyber knife.   Standard CRT techniques generate highly conformal radiotherapy 
  using radiation fields generated in a few fixed-angle positions. These are 
  generated by linear accelerators (aka. LINAC) which are then modulated by 
  conventional leaf collimators operated in either dynamic mode or the standard 
  step-and-shoot mode. CyberKnife is a modern radiotherapy system developed by 
  Accuray, Inc., primarily for radiosurgery. It makes use of a highly precise 
  robotic arm to delivery the doses. The arm is computer driven to achieve very 
  high precision, it is mounted with a compact 6-MV LINAC whose orientation is 
  determined by 6 different mechanical subsystems. This robotic arm can be 
  moved with a much greater generality than most conventional sytsems.

  Table \ref{tab:compare} highlights the key differences in various systems. 
  For the 
  sake the completeness, the authors of Ref. \cite{Fenwick2006} have 
  dicotomized the conventional system into 2 types, with and without a 
  multileaf collimator, and the tomotherapy into serial and parallel 
  tomotherapy.
  \section{Clinical Applications}
  \label{clinical}
  This section highlights few of the many clinical applications of tomotherapy. 
  \subsection{Total Body Irradiation (TBI)}
  Tomotherapy can be used in establishing total body irradiation (TBI) to gain 
  greater control over the dose distribution and to spare organs that may be at 
  risk. It was shown in Ref. \cite{Gruen2013} that using HT provided excellent 
  conformal 
  lung sparing with mean doses not exceeding 10 Gy. No lung toxicity was 
  observed for over 15 months after follow-up. Also, the overall morbidity 
  during the treatment period was very low, corresponding to grades 1-2 side 
  effects.

@@

  \subsection{Whole brain helical Tomotherapy}
  \subsection{Inoperable Lung Cancer}
  \subsection{Rectal Cancer}
  \section{Conclusion}
  \label{conclusions}
  \bibliography{Paper}{}
  \bibliographystyle{IEEEtran}

>>>>

\documentclass[10pt,journal,compsoc]{IEEEtran} %draftclsnofoot, 
\usepackage{graphicx}
\graphicspath{ {./_figures/} }
\setlength\fboxsep{1pt}
\setlength\fboxrule{1pt}
\usepackage{multicol}
\bstctlcite{IEEEexample:BSTcontrol}
\usepackage{bigstrut}
\usepackage{tabularx}
\usepackage{tabulary}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage{balance}
\usepackage{flushend}
\setlength{\parindent}{0em}
\setlength{\parskip}{1em}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{hyperref}
\hypersetup{
  colorlinks = false,
  hidelinks = true
	}
\usepackage[english]{babel}
\usepackage{blindtext}
\usepackage{times}
\usepackage{cite}

@@

  The setup has a 
  linear accelerator mounted on a rotating gantry as seen in CT 
  sacnners. The radiation is delivered to a patient in a helical way, obtained 
  by concurrent gantry rotation and couch/patient travel as shown in Fig. 
  \ref{fig1}. In general, the helical tomotherapy units have significantly 
  different radiation fields when compared to other radiotherapy systems. This 
  difference can be owed to the absence of a flattening filter, a 
  thin target, an electron stopper, a beam hardener and a compact primary 
  collimator. This paper will briefly highlight the 
  radiation characteristics of the helical tomotherapy unit, especially in 
  parts where it differs from standard accelerators.

@@

      \toprule[0.4mm]

@@

      \midrule[0.4mm]

@@

  Figure \ref{radcharb} shows the differences in the photon spectrum between 
  the 
  treatment and imaging modes of helical tomotherapy. Notice that the in the 
  imaging mode the spectrum has a lower energy when compared to treatment mode; 
  this is done so as to obatin better images by improving the contrast. A more 
  comprehensive comparision is documented in Ref. \cite{Jeraj2004}.

@@

      \label{tab:compare}%
    \end{table*}% \be
  Tomotherapy is different from other standard IMRT techiniques such as 
  conformal radiotherapy (CRT) techniques and non-standard IMRT techniques like 
  Cyber knife.   Standard CRT techniques generate highly conformal radiotherapy 
  using radiation fields generated in a few fixed-angle positions. These are 
  generated by linear accelerators (aka. LINAC) which are then modulated by 
  conventional leaf collimators operated in either dynamic mode or the standard 
  step-and-shoot mode. CyberKnife is a modern radiotherapy system developed by 
  Accuray, Inc., primarily for radiosurgery. It makes use of a highly precise 
  robotic arm to delivery the doses. The arm is computer driven to achieve very 
  high precision, it is mounted with a compact 6-MV LINAC whose orientation is 
  determined by 6 different mechanical subsystems. This robotic arm can be 
  moved with a much greater generality than most conventional sytsems.

  Table \ref{tab:compare} highlights the key differences in various systems. 
  For the 
  sake the completeness, the authors of Ref. \cite{Fenwick2006} have 
  dicotomized the conventional system into 2 types, with and without a 
  multileaf collimator, and the tomotherapy into serial and parallel 
  tomotherapy.
  \section{Clinical Applications}
  \label{clinical}
  This section highlights few of the many clinical applications of tomotherapy. 
  \subsection{Total Body Irradiation (TBI)}
  Tomotherapy can be used in establishing total body irradiation (TBI) to gain 
  greater control over the dose distribution and to spare organs that may be at 
  risk. In a study conducted by Ref. \cite{Gruen2013}, with a cohort of 10 
  patients aged 4 - 22 years with acute lymphoblastic or myeloic leukemia, it 
  was shown that, among other benifits, using HT provided excellent 
  conformal lung sparing with mean doses not exceeding 10 Gy. No lung toxicity 
  was observed for over 15 months after follow-up. Also, the overall 
  morbidity   during the treatment period was very low, corresponding to grades 
  1-2 side   effects. Table \ref{tbi_side} from Ref. \cite{Gruen2013} documents 
  other side  effects, notice that none of the side effects for any patient 
  were of grades 3 or 
  4.

  \begin{table}[t]
    \caption{Acute Toxicity (Ref. \cite{Gruen2013})}
    \centering
    \includegraphics[width=\linewidth]{fig8.png}
    \label{tbi_side}
  \end{table}

@@

  \subsection{Whole brain helical Tomotherapy}
  Whole brain helical tomotherapy is a possible treatment option for patients 
  suffering from malignant melanoma with four or more brain metastates. The 
  possiblity of integrated boost with enhanced hippocampal sparing makes the 
  use of HT a very reliable approach.

  A study conducted by Ref. \cite{Gondi2010} showed that Helical 
  tomotherapy 
  spared the hippocampus, with a median dose of 5.5 Gy and maximum 
  dose of 12.8 
  Gy. Other linear accelerator based IMRT spared the hippocampus, 
  with a median 
  dose of 7.8 Gy and maximum dose of 15.3 Gy. On a per-fraction 
  basis, mean 
  dose to the hippocampus (normalized to 2-Gy fractions) was reduced 
  by 87\% to 
  0.49 Gy2 using helical tomotherapy and by 81\% to 0.73 Gy2 using 
  LINAC-based 
  IMRT. Target coverage and homogeneity was acceptable with both IMRT 
  modalities, with differences largely attributed to more rapid dose 
  falloff 
  with helical tomotherapy.

  Another study was conducted by Ref. \cite{Hauswald2013} to explore 
  and 
  compare the 
  effects of treatment and response to conventional radiotherapy with 
  a 30 Gy 
  dose in 10 fractions and a standard helical tomography with 30 Gy 
  dose, 10 
  fraction, integrated boost of 50 Gy at the regions of metastates 
  with a 
  hippocampal-sparing. This study showed that the even though HT 
  produced 
  excellent homogeneity and conformality with integrated boost, it 
  still has 
  some limitations when compared to classical radiotherapy on account 
  of the 
  fact that all HT plans had a higher dose brain exposure. The 
  effects 
  shouldn't be too much of an issue as the total treatment time is 
  lower, and 
  so too the integral dose, due to the dynamic couch setup and other 
  design 
  aspects of the HT system.
  \subsection{Inoperable Lung Cancer}
  Radiation therapy has become a standard treatment protocol for 
  patients 
  suffering from non-small cell lung cancer and other solitary lung 
  metastatis. 
  The use had been quite controversial. Among other things, the 
  primary 
  concerns were the expansion of the low-dose irradiated lung volume 
  resulting 
  from the coplanar beam delivery and uncertainity in dose 
  distribution due to 
  respiration.

  However, a very recent study published by Ref. \cite{Nagai2014} seem 
  to show 
  that Helical Tomotherapy for lung cancer is both safe and 
  effective. In this 
  study 2 groups of lung tumor patients were treated; one group of 37 
  patients 
  were administered a dose of 48 Gy in 4 fractions and another group 
  of 35 
  patients were administered a dose of 50-60 Gy in 5-8 fractions. The 
  4-fraction group had a median planning target volume (PTV) of 6.9 
  cm\textsuperscript{3} and the 5-8 fraction group had a median PTV 
  of 14 
  cm\textsuperscript{3}. The study reports that only 2 of the 37 and 
  2 out of 
  35 patients in the 4-fraction cohort respectively had Grade-2 
  radiation 
  pneumonitis. No other complications were reported.

  \subsection{Colorectal Cancer}
  A study published by Ref. \cite{Yu2013} showed that using 
  tomotherapy 
  could produce acceptable coverage for the target while decreasing 
  the mean dose to sensitive surrounding areas such as bowel, bladder 
  and femur heads. They report that while tomotherapy had a  lower 
  radical 
  dose homogeneity index (rHDI) compared to conventional 3D conformal 
  radiotherapy (3D-CRT) (0.7$\pm$0.01 vs 22.05$\pm$2.76, p$<$0.001),  
  it did significantly reduced the volume of sensitive tissues 
  receiving a large dose; this was measured using $V_{nGy}$, which 
  denotes the percentage of volume receiving a dose greater than n 
  Gy. Figure \ref{fig6}, taken from \cite{Yu2013}, shows the dose 
  reduction 
  acheived by tomotherapy.
  \begin{figure}[t]
		\centering
    \includegraphics[width=\linewidth]{fig9.png}
    \caption{Tomotherapy(dashed blue line) significantly reduce 
    $V_{nGy}$ for the bowel as compared to 3D-CRT (with p<0.001) for 
    all doses except at 5 Gy (p = 0.568), Ref. \cite{Yu2013}.}
  	\label{fig6}
  \end{figure}

  \subsection{Arteriovenous Malformations}
  Arteriovenous malformations (AVM) are aberrant fistulous connections between 
  arteries and veins bypassing the normal capillary bed. It occurs in 0.2-0.8\% 
  of the normal population. If untreated cerebral AVMs can cause hemorrhage and 
  subsequent neurological impairments. An attempt was made by Ref. 
  \cite{Krause2013} to 
  validate the performance of modern tomotherapy systems in tackling AVM.

  They 
  reported that there were no significant results to show that tomotherapy 
  outperforms the conventional linac based methods. In fact, as previously 
  reported by many researchers, tomotherapy exposed bigger brain volumes 
  to doses $\ge60\%$ of the prescription dose (average 12.75 Gy), albeit only 
  for a short amount of time. This does, unfortunately, pose risks of 
  permanent postradiosurgery sequelae influenced by the localisation and the 
  volume of brain receiving 12 Gy or more. However, these were only theoretical 
  planning comparison to explore the potential of these methods for small 
  volume radiosurgery. HT radiosurgery might yet be beneficial for AVM with 
  special geometry as well as for patients with brain metastases that benefit 
  from shorter treatment times at the cost of a slightly higher brain exposure.

  \section{Conclusion}
  \label{conclusions}
  In conclusion, IMRT with helical tomotherapy is an advanced form of 
  radiotherapy which constantly outperforms several conventional 3D-CRT. It 
  represents one of the most important technical advances in radiation therapy 
  since the advent of medical linear accelerator.

  Current helical tomography systems provide excellent target volume 
  conformality while significantly reduction exposure to the surrounding 
  sensitive tissue. This aspect of the tomotherapy system has motivated its use 
  in oncology. Some clincal applications are listed above, this list by no 
  means exhaustive. However, it needs to be mentioned that the results are 
  reported for small cohorts and in many cases reports are only feasibility 
  studies. There is a lot active research in the field and as Ref. 
  \cite{Beavis2004} 
  suggests, helical tomotherapy can be \textit{a} future of IMRT if not 
  \textit{the} future.

  \bibliography{Paper}{}
  \bibliographystyle{IEEEtran}
